Symbolic Abstract Data Type Inference

### 2015-07-09 inference.tex
Reference to proof.

Let $H^*$ denote the \emph{closure} of a history set $H$ under the relation
$\mathord\to = (\to_\mathrm{o} \cup \to_\mathrm{p} \cup \to_\mathrm{c})$, i.e.,
\begin{align*}
  H^* = \set{ h' : \exists h \in H.\ h \to^\ast h' }.
\end{align*}
A set $K$ \emph{generates} $H$ when $K^* = H$, and a \emph{kernel} of a set $H$
is a minimal set $K$ which generates $H$. While the kernels of arbitrary sets of
histories need not be unique, the kernels of sets which have sequential kernels
are unique. Furthermore, Section~\ref{sec:nature} demonstrates that the
histories of naturally-occurring implementations have unique kernels, which we
assume for the remainder of this work. An \emph{abstract data type (ADT)} $A$ is
the kernel of the set $H(\mathcal{I})$  of histories of some implementation
$\mathcal{I}$. We say that \emph{$A$ is the ADT of $\mathcal{I}$} when $A^* =
H(\mathcal{I})$.

>>>>

A Set $K$ \emph{generates} $H$ when the closure of $K$ under the relation
$\mathord\to = (\to_\mathrm{o} \cup \to_\mathrm{p} \cup \to_\mathrm{c})$
is equal to $H$, i.e., $H = \set{ h : \exists h' \in K.\ h' \to^\ast h }$.
A \emph{kernel} of a set $H$ is a minimal set generating $H$. While the kernels
of arbitrary sets need not be unique, the kernels of sets which
have sequential kernels are unique. Furthermore, Section~\ref{sec:nature}
demonstrates that the histories of naturally-occurring implementations have
unique kernels, which we assume for the remainder of this work.
An \emph{abstract data type (ADT)} $A$ is the kernel of the set $H(\mathcal{I})$
of histories of some implementation $\mathcal{I}$; we say that \emph{$A$ is the
ADT of $\mathcal{I}$}.

### 2015-07-09 patterns.tex
Reference to proof.

The previous example demonstrates that $\preceq$ is not a wqo by allowing each
history $h_i$ of the infinite sequence to contain more and more pending
operations in order to ensure that $h_j \not\preceq h_i$ for every $j < i$.
Curbing this ability by limiting the maximum amount of pending operations per
history makes $\preceq$ a wqo. Although limiting to $k$ pending operations
essentially limits us to width-$k$ histories, of executions with at most $k$
operations parallel at any moment, e.g.,~of programs with at most $k$ threads,
this restriction comes at no loss of completeness when considering only the
histories of bounded-width ADTs, which is the subject of the remainder
of this section.

@@

  $\preceq$ is a wqo on width-bounded histories.

@@

For the remainder of this section, we fix an ADT $A$, and let $B$ be its bounded
complement. When $A$ has bounded width, so does $B$, and thus $\preceq$ is a wqo
on $B$. Furthermore, when $A$ normal, it is closed under $\succeq$, and thus $B$
is closed under $\preceq$~\footnote{Actually, $B$ is closed under $\preceq$ when
restricted to width-bounded histories, i.e., if $h_1 \preceq h_2$, $h_1 \in B$,
and $h_2$ is width-bounded, then $h_2\in B$.}. Closure under a relation
satisfying Lemma~\ref{lem:wqo} implies representation by a finite set. Formally,
we say a set $X$ is \emph{finitely representable} if there exists a finite set
$Y$ and a relation $R \subseteq Y \times X$ such that $X = \set{ x : \exists
y\in Y.\ R(y,x) }$. In our case, we obtain a finite set from which exactly the
elements of $B$ are related by $\preceq$.

>>>>

This example demonstrates that $\preceq$ is not a wqo by allowing each history
$h_i$ of the infinite sequence to contain more and more pending operations in
order to ensure that $h_j \not\preceq h_i$ for every $j < i$. Curbing this
ability by limiting the maximum amount of pending operations per history makes
$\preceq$ a wqo. Although limiting to $k$ pending operations essentially limits
us to width-$k$ histories, of executions with at most $k$ operations parallel at
any moment, e.g.,~of programs with at most $k$ threads, this restriction comes
at no loss of completeness when considering only the histories of bounded-width
ADTs, which is the subject of the remainder of this section.

@@

  $\preceq$ is a wqo on bounded-width histories.\footnote{The proof of
  Lemma~\ref{lem:wqo} appears in Appendix~\ref{sec:proof}.}

@@

For the remainder of this section, we fix an ADT $A$, and let $B$ be its bounded
complement. When $A$ has bounded width, so does $B$, and thus $\preceq$ is a wqo
on $B$. Furthermore, when $A$ normal, it is closed under $\succeq$, and thus $B$
is closed\footnote{Actually, $B$ is closed under $\preceq$ when restricted to
width-bounded histories, i.e., if $h_1 \preceq h_2$, $h_1 \in B$, and $h_2$ is
width-bounded, then $h_2\in B$.} under $\preceq$. Closure under a relation
satisfying Lemma~\ref{lem:wqo} implies representation by a finite set. Formally,
we say a set $X$ is \emph{finitely representable} if there exists a finite set
$Y$ and a relation $R \subseteq Y \times X$ such that $X = \set{ x : \exists
y\in Y.\ R(y,x) }$. In our case, we obtain a finite set from which exactly the
elements of $B$ are related by $\preceq$.
